---
title: "Overview of the capstoneJH package"
author: "Dimitris Gateley"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Overview of the capstoneJH package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, include = FALSE}
library(capstoneJH)
library(dplyr)
library(ggplot2)
library(readr)
```
## Package functions

There are six exported functions available to users:

- `eq_clean_data()`
- `geom_timeline()`
- `geom_timeline_label()`
- `theme_timeline()`
- `eq_create_label()`
- `eq_map()`

Data from NOAA can be found in the package directory under `\extdata` folder.

## Clean data

The first function is required to clean data for the visualization. It creates a DATE column in `Date` format, transforms latitude and longitude to numeric format and trims country from LOCATION_NAME.

```{r eq_read_example, message = FALSE}
filename <- system.file("extdata/earthquakes.tsv.gz", package = "capstoneJH")
data <- readr::read_delim(filename, delim = "\t")

eq_clean_data(data)
```

The `geom_timeline()` requires clean data.

```{r eq_timeline_example, fig.width = 7, fig.height = 4}
data %>% eq_clean_data() %>%
     filter(COUNTRY %in% c("GREECE", "ITALY"), YEAR > 2000) %>%
     ggplot(aes(x = DATE,
                y = COUNTRY,
                color = as.numeric(TOTAL_DEATHS),
                size = as.numeric(EQ_PRIMARY)
     )) +
     geom_timeline() +
     geom_timeline_label(aes(label = LOCATION_NAME), n_max = 5) +
     theme_timeline() +
     labs(size = "Richter scale value", color = "# deaths")
```

## Visualize earthquakes on map

The package utilized `leaflet` functions to visualize earthquakes on a map using `eq_map()` function. The map is automatically trimmed to display the input data frame. Optional annotations can be created using `eq_create_label()` function. The result is an interactive map where user can click on individual points to get details:

```{r eq_map_example, fig.width = 7, fig.height = 4}
data %>% 
  eq_clean_data() %>% 
  dplyr::filter(COUNTRY == "GREECE" & lubridate::year(DATE) >= 1990) %>% 
  dplyr::mutate(popup_text = eq_create_label(.)) %>% 
  eq_map(annot_col = "popup_text")
```
